CODECHECK: An open-science initiative for the independent execution of computations underlying research articles during peer review improves reproducibility

### 2021-02-01 codecheck-paper.tex
small edits by DN before revision

CODECHECK is best demonstrated by way of our example workflow, and later
we expand on the underlying principles. The workflow involves three
groups of people:
(1) the AUTHOR of a paper providing the code to be checked,
(2) the PUBLISHER of a journal interested in publishing the paper 
by the author, and
(3) the CODECHECKER who checks that the authors's code works.
The workflow that we have refined is documented in Figure~\ref{fig:workflow}
and consists of six steps.

@@

\textbf{Step 6:} the PUBLISHER can, depending on the timing, provide the
certificate to scientific reviewers or handling editors, or publish it
and link between the certificate, the paper, and any repositories 
through metadata. Currently, the CODECHECKER creates these on Zenodo\footnote{
See ``related identifiers'' and ``alternate identifiers'' in the right-hand
sidebar of certificate \texttt{2020-025} \cite{cert-2020-025}.}.
The PUBLISHER also gives credit to the CODECHECKER for their work by depositing
the activity appropriately in scholarly profiles, such as ORCID\footnote{
See peer review contributions in ORCID records at
\url{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}.},
The PUBLISHER also ensures proper publication metadata, e.g.,~links from the 
certificate repository to the published paper or the original code repository.

@@

In contrast, for researchers it can be very important to be publicly credited
with their activity as a reviewer.
A regular
review may be listed in public databases (e.g.,~ORCID
\footnote{\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}} or commercial offerings such as Publons
\footnote{\href{https://publons.com/}{https://publons.com/}} or 
ReviewerCredits\footnote{\href{https://www.reviewercredits.com/}{https://www.reviewercredits.com/}}).
A codechecker should be listed just the same. The number of scientific reviews
is of course a very coarse indicator for the community contribution by an 
individual, yet it is unclear if a regular reviewer, who in addition
conducts a CODECHECK to the scientific review, should be credited twice.

@@

\textbf{3. Credit is given to codecheckers.} The value of contributing a
CODECHECK is comparable to providing a scientific peer review. Individual
variations aside, it may also require a similar amount of time. Therefore,
the codechecker's activity should be publicly recorded and mentioned in the
published paper.
The public record can be realised by publishing the certificate in a 
citable form (i.e.,~with a DOI; this is done within the community codecheck
process), by listing codecheckers on the journal's website or, ideally, by
publishing the checks alongside peer review activities in public databases.

@@

The third theme represents DN's service as Reproducibility
Reviewer at the AGILE conference series.
Independent of CODECHECK, the \emph{Reproducible AGILE} Initiative \cite{reproducible_agile}
established a process for reproducing workflows at the AGILE conference
series \cite{nust_improving_2020}.
While using slightly different terms and infrastructure (``reproducibility
reports'' are published on OSF instead of certificates on Zenodo) the 
AGILE reproducibility reviews adhere to the codecheck principles.
Furthermore, a small number of checks was completed as part of a peer reviews
for GIScience journals.

@@

To run the system, we rely on freely available infrastructure, more 
specifically GitHub and Zenodo.

The \texttt{codecheckers} \textbf{GitHub} organisation\footnote{
\url{https://github.com/codecheckers}} contains projects for managing the
project website, the community of codecheckers, community discussions,
forks of checked code repositories, and the main register of 
CODECHECKs. Both the website\footnote{
\url{https://codecheck.org.uk/}} and the register\footnote{
\url{https://codecheck.org.uk/register}} are hosted as GitHub
pages. The register is a single table in CSV format that
connects the certificate identifier with the repository associated with
a specific conducted CODECHECK. Each of these repositories, which 
currently can be hosted on GitHub or Open Science Framework (OSF), 
contains the CODECHECK metadata file \texttt{codecheck.yml}. The register
further contains a column for the type of check, e.g.,~community, journal,
or conference, and the respective GitHub issue where communications and 
assignments around a specific check are organised. No information is 
duplicated between the register and the metadata files. When a new
certificate is entered into the register, an automated process using the
continuous integration infrastructure of GitHub, GitHub Actions, retrieves
all metadata files and builds different representations of the register
for viewing (HTML) and for integration with other services (CSV, JSON).

\textbf{Zenodo} is an open repository for scientific data and documents. It
mints DOIs for deposits and ensures long-term availability of all digital
artefacts related to the project. These include the community 
certificates as well as the regularly archived register \cite{codecheck_register}.

A \textbf{custom R package}, \texttt{codecheck}\footnote{
\url{https://github.com/codecheckers/codecheck}}, helps with the authoring
of certificates, and with the deposition of certificates and realed files to
Zenodo using the R package \texttt{zen4R} \cite{zen4r}.
Codecheckers may choose to not use the package and rely on their own tools
for creating the certificate, even a simple word processor.
This is a deliberate flexibility to
accomodate different skill sets and unforseen technical advances or
challenges.

@@

\textbf{CODECHECK requirements are not demanding enough}:
CODECHECK by design does not require authors to provide (and sustain) an
eternally functional workflow and neither suggests a specific software stack
or practical approaches, such as research compendia
\footnote{\url{https://research-compendium.science/}},
even if they are very valuable.
Creating something that anyone can reproduce has been called a 
fool's errand
\footnote{\url{https://twitter.com/DougBlank/status/1135904909663068165?s=09}}
and we tend to agree.
However, the package of data, code, and documentation
collaboratively created by authors and codecheckers is a snapshot of a 
working analysis that greatly increases the likelihood of a successful 
reproduction and possibility to extensions by third parties, with access
to suitable resources and matching skill set, in the future.
Concrete implementations of CODECHECK workflows, especially for specific
disciplines, may reify much more helpful guidelines for authors how to
create reproducibility packages.
The author-friendly ``low bar'' should not stay low forever, but cultural
change takes time and the encouragement and guidance that CODECHECK,
as part of the widely accepted peer review concept, can provide may
eventually allow to raise the bar as high as possible, e.g.,~with
executable research compendia \cite{nust_opening_2017}
or continuous analysis \cite{beaulieu-jones_reproducibility_2017-1}.
However, considering that missing artefacts and lack of documentation
have repeatedly been identified as main barriers to reproducibility
(e.g.,~\cite{stagge_assessing_2019,nust_improving_2020}),
we would not underestimate the power of a simple check.
ModelDB curation policies ensure manually to reproduce one figure only 
\cite{mcdougal_reproducibility_2016}, but that has not limited the usefulness
nor success of the platform.

@@

An important distinction is that a codechecker does not conduct a 
\emph{code review}. Code reviews are without doubt extremely valuable in 
terms of improving reproducibility and re-usability, and their proponents 
even believe they can improve the science \cite{petre_code_2014}.
Code reviews, however, have quite different structural challenges and 
require even more resources. That said, a well-reviewed codebase is likely
to be easier to codecheck, and the awareness for high-quality code raised
through CODECHECK may lead to more support for code reviewing.
Initiatives and journals that conduct software reviews independently of 
a specific publication or workflow include
ROpenSci\footnote{\url{https://ropensci.org/}},
PyOpenSci\footnote{\url{https://www.pyopensci.org/}},
and JOSS\footnote{\url{https://joss.theoj.org/}}.

Furthermore, the codechecker's task list is intentionally not
overloaded with related concerns, even though they are very important,
such as ensuring proper citation of data and software, or depositing
data and software in suitable repositories. We do expect codecheckers
to be aware of these though and possibly highlight them.

\textbf{Handling failure}: We have no established process for the case
that a reproduction fails. This may be the sign of a pre-selection bias
of the examples, as so far all our CODECHECKs succeeded.
In case of a journal adopting CODECHECK for all submissions, the open 
question remains as what to do in this case will have to be answered.
At the start, we doubt publicly reporting failures (i.e.,~the code 
would not run) will increase overall reproducibility and recommend not
to share the outcome publicly but to share the latest draft of the 
certificate with the authors.
A negative comment in a CODECHECK certificate or a failed check does
not necessarily mean the paper or the science is bad
(cf. discussion on negative comments in \cite{everythinghertz123}).
Therefore, one must be careful as to the impact of acceptance by authors 
and the concerns of volunteers this approach, or a contrary one, has.
Private feedback for failures and public acknowledgement/encouragement
for successes seems to be the way to go.
% eLife's new model (see https://osf.io/9cftx/) will publish reports evantually, e.g.,~after preprint has been accepted elsewhere
\cite{Rosenthal2016b} discuss incentives for different actors around the
implementation of reproducibility. We see CODECHECK as one mean for
organisations to invest in reproducibility by creating incentives
until reproducible computations become the norm.

\textbf{Compute time}: for those papers that take significant compute
time (think days, not minutes), who will pay for the compute time?
The openness of CODECHECK certificates could help to establish community
peer pressure, as submitting authors can be evaluated on their engagement
as codecheckers.
The challenges of computational resources can also quickly be reduced when
authors provide a small sample or a synthetic dataset together with the
expected outputs for this small example, which can equally well demonstrate the 
usefulness and validity of their workflow.

\textbf{Proprietary software}: authors can currently provide code that requires
proprietary software. Given the prevalence of proprietary software in
some disciplines, e.g MATLAB, this seems like a pragmatic choice.
However, it prohibits to benefit from open infrastructure for reproducibility
(cf.~\cite{konkol_publishing_2020,perkel_make_2019})
and requires the codechecker to have access to that software.
Non-open software also considerably hampers reuse, especially by researchers
from the global south. Therefore, allowing proprietary software is a compromise
that each implementation of CODECHECK should critically reconsider.

\textbf{Not possible for my data}: 
Solutions for proprietary and sensitive data exists.
Authors can provide synthetic data (cf. \cite{shannon_opening_2018}), some
data can effectively be redacted \cite{oloughlin_data_2015}, and publishers or
independent entities can provide infrastructure for sharing data and workflows
confidentially \cite{perignon_certify_2019} or with access to derived results
but not raw data \cite{shannon_opening_2018},
i.e.,~data enclaves \cite{foster_research_2018},
or domains of reproducibility \cite{harris_more_2017})

@@

\textbf{Who's got time for more peer review?} Agree.
Ignoring the complex community contract behind volunteered peer review,
the time spent with CODECHECKs likely reduces the individual's time to work
as a reviewer.
However, there are differences to peer review.
First, the technical nature of a CODECHECK can provide a more 
clearly defined scope and expectations compared to conventional 
peer review. The authors are told what to provide and the codechecker
can be told when to stop, for example.
Second, the specific skill set allows to introduce different groups
into the peer review process. ECRs might be attracted to working with
groups to learn more about recent methods and reproducibility practices.
Research software experts (Research Software Engineers, RSE) who might
not regularly be involved in writing or reviewing papers could be 
interested to increase their connection with scholarly practices.
An extra codechecker may simplify the matchmaking an editor has to do
when identifying suitable reviewers for a submission, as technical and
topical expertise can be provided by different people, who must not even
be traditional peer reviewers. In fact, the codechecker could even be
a software professional employed by a publisher.
Third, recall that we expect CODECHECKs to be not anonymous and always
publicly deposited, which is common practice for some journals, but not
broadly practiced. The possibility to ask for clarification or more 
material/information can greatly reduce the digging required and thereby
the time spend.
We also found that the focus on the mechanics of a
workflow and the ability to interact with the authors can make
reproductions very educational. It's a different role and as such could
be a welcome alternation/diversion for researchers to give back their 
time to the community.
With code and workflows, there also is a good chance that a codechecker's
feedback can have direct impact and lead to improvements of the author's 
work that comes with a rewarding experience or even tangible contributions.
While such impact is also part of idealistic peer review, it is
rather rare beyond paraphrased anonymous acknowledgement.

@@

As a principle, we do not require results to be exactly the same for a
CODECHECK to pass, simply that the code runs and generates the output
files that the author claims. Stochastic simulations mean that often
we cannot get exactly the same, or even the same versions of libraries
can generate outputs that differ by operating system
\cite{Gronenschild2012-pp}.

\textbf{Not revolutionary enough?}
CODECHECK's approach is to acknowledge the shortcomings around 
computational reproducibility and iteratively improve within the current
system of scholarly communication through journals.
It remains to be proved if this approach is welcomed broadly and if 
involving the traditional system of publishers helps to push the 
conversation along or not.
We have discussed stronger more drastic rules at length, such as only 
considering diamond Open Access journals, community-led no-fee journals,
or journals with collaborative/open peer review, but eventually decided
against them.
We also have considered to require latest technologial solutions which can
support reproducibility (cf.~\cite{konkol_publishing_2020}), but have 
decided to focus on the human interface and the judgement of an experienced
researchers/developer as a more sustainable approach.
The structurelessness of the required data and software gives the needed
freedom to enable all types of research. It could be complemented by
automated scoring (e.g.,~\cite{menke_rigor_2020}, yet automation and 
metrics bears risks that lead us to put our trust in the codechecker.
Note that the focused nature of the CODECHECK principles allows journals
and publishers to innovate on the financial model and peer review practices
at their own pace.

>>>>

CODECHECK is best demonstrated by way of our example workflow, and later
we expand on the underlying principles. The workflow involves three
groups of people:
(1) the AUTHOR of a paper providing the code to be checked,
(2) the PUBLISHER of a journal interested in publishing the paper 
by the author, and
The workflow we have refined is shown in Figure~\ref{fig:workflow}
and consists of six steps.

@@

\textbf{Step 6:} the PUBLISHER can, depending on the timing, provide the
certificate to scientific reviewers or handling editors, or publish it
and link between the certificate, the paper, and any repositories 
through metadata. Currently, the CODECHECKER creates these connections on
Zenodo \footnote{
See ``related identifiers'' and ``alternate identifiers'' in the right-hand
sidebar of certificate \texttt{2020-025} \cite{cert-2020-025}.}.
The PUBLISHER also gives credit to the CODECHECKER for their work by depositing
the activity appropriately in scholarly profiles, such as ORCID\footnote{
See peer review contributions in ORCID records at
\url{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}.}.
The PUBLISHER also ensures proper publication metadata, e.g.,~links from the 
certificate repository to the published paper or the original code repository.

@@

In contrast, for researchers it can be very important to be publicly credited
with their activity as a reviewer.
A regular
review may be listed in public databases (e.g.,~ORCID
\footnote{\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}} or commercial offerings such as Publons
\footnote{\href{https://publons.com/}{https://publons.com/}} or 
ReviewerCredits\footnote{\href{https://www.reviewercredits.com/}{https://www.reviewercredits.com/}}).
A codechecker should be listed just the same. The number of scientific reviews
is of course a very coarse indicator for the community contribution by an 
In this case, the published CODECHECK certificate and possibly the published peer review can clearly describe the possibly large efforts put into both.
conducts a CODECHECK to the scientific review, should be credited twice.

@@

\textbf{3. Credit is given to codecheckers.} The value of contributing a
CODECHECK is comparable to providing a scientific peer review. Individual
variations aside, it may also require a similar amount of time. Therefore,
the codechecker's activity should be publicly recorded and mentioned in the
published paper.
The public record can be realised by publishing the certificate in a 
citable form (i.e.,~with a DOI, as conducted in the community CODECHECK
process), by listing codecheckers on the journal's website or, ideally, by
publishing the checks alongside peer review activities in public databases.

@@

The third theme represents DN's service as a Reproducibility Reviewer at the AGILE conference series, where the \emph{Reproducible AGILE} Initiative \cite{reproducible_agile} established a process for reproducing workflows at the AGILE conference series independent of CODECHECK \cite{nust_improving_2020}.
While using slightly different terms and infrastructure (``reproducibility
reports'' are published on the Open Science Framework (OSF) instead of certificates on Zenodo) the 
AGILE reproducibility reviews adhere to the CODECHECK principles.
Furthermore, a small number of checks was completed as part of a peer reviews
for GIScience journals.

@@

To run the system, we rely on freely available infrastructure, more 
specifically GitHub and Zenodo.
The \texttt{codecheckers} \textbf{GitHub} organisation\footnote{
\url{https://github.com/codecheckers}} contains projects for managing the
project website, the community of codecheckers, community discussions,
forks of checked code repositories, and the main register of 
CODECHECKs. Both the website\footnote{
\url{https://codecheck.org.uk/}} and the register\footnote{
\url{https://codecheck.org.uk/register}} are hosted as GitHub
pages. The register is a single table in CSV format that
connects the certificate identifier with the repository associated with
a specific conducted CODECHECK. Each of these repositories, which 
currently can be hosted on GitHub or Open Science Framework (OSF), 
contains the CODECHECK metadata file \texttt{codecheck.yml}. The register
further contains a column for the type of check, e.g.,~community, journal,
or conference, and the respective GitHub issue where communications and 
assignments around a specific check are organised. No information is 
duplicated between the register and the metadata files. When a new
certificate is entered into the register, an automated process using the
continuous integration infrastructure of GitHub, GitHub Actions, retrieves
all metadata files and builds different representations of the register
for viewing (HTML) and for integration with other services (CSV, JSON).
\textbf{Zenodo} is an open repository for scientific data and documents. It
mints DOIs for deposits and ensures long-term availability of all digital
artefacts related to the project. The CODECHECK community on Zenodo
\footnote{\url{https://zenodo.org/communities/codecheck/}} includes 
certificates as well as the regularly archived register
\cite{codecheck_register_jan2021}.

A \textbf{custom R package}, \texttt{codecheck}\footnote{
\url{https://github.com/codecheckers/codecheck}}, is used to automate repetitive
tasks around the authoring of certificates and the management of the register.
For example, it includes scripts for the deposition of certificates and related 
files to Zenodo using the R package \texttt{zen4R} \cite{zen4r} and for the
register update process outlined above.
Codecheckers may choose to not use the package and rely on their own tools
for creating the certificate, even a regular word processor.
This is a deliberate flexibility to
accomodate different skill sets and unforseen technical advances or
challenges.

@@

\textbf{Aren't CODECHECK requirements not demanding enough?}
CODECHECK by design does not require authors to provide (and sustain) an
eternally functional workflow and neither suggests a specific software stack
or practical approaches, such as research compendia
\footnote{\url{https://research-compendium.science/}},
even if they are very valuable.
Creating something that anyone can reproduce has been called a 
fool's errand
\footnote{\url{https://twitter.com/DougBlank/status/1135904909663068165?s=09}}
and we tend to agree.
However, the package of data, code, and documentation
collaboratively created by authors and codecheckers is a snapshot of a 
working analysis that greatly increases the likelihood of a successful 
reproduction and possibility to extensions by third parties, with access
to suitable resources and matching skill set, in the future.
Concrete implementations of CODECHECK workflows, especially for specific
disciplines, may reify much more helpful guidelines for authors how to
create reproducibility packages.
The author-friendly ``low bar'' should not stay low forever, but cultural
change takes time and the encouragement and guidance that CODECHECK,
as part of the widely accepted peer review concept, can provide may
eventually allow to raise the bar as high as possible, e.g.,~with
executable research compendia \cite{nust_opening_2017}
or continuous analysis \cite{beaulieu-jones_reproducibility_2017-1}.
However, considering that missing artefacts and lack of documentation
have repeatedly been identified as main barriers to reproducibility
(e.g.,~\cite{stagge_assessing_2019,nust_improving_2020}),
we would not underestimate the power of a simple check.
ModelDB curation policies ensure manually to reproduce one figure only 
\cite{mcdougal_reproducibility_2016}, but that has not limited the usefulness
nor success of the platform.

@@

An important distinction is that a codechecker does not conduct a 
\emph{code review}. Code reviews are without doubt extremely valuable in 
terms of improving reproducibility and re-usability, and their proponents 
even believe they can improve the science \cite{petre_code_2014}.
Code reviews, however, have quite different structural challenges and 
require even more resources. That said, a well-reviewed codebase is likely
to be easier to codecheck, and the awareness for high-quality code raised
through CODECHECK may lead to more support for code reviewing.
Initiatives and journals that conduct software reviews independently of 
a specific publication or workflow include
ROpenSci\footnote{\url{https://ropensci.org/}},
PyOpenSci\footnote{\url{https://www.pyopensci.org/}},
and JOSS\footnote{\url{https://joss.theoj.org/}}.
Furthermore, the codechecker's task list is intentionally not
overloaded with related concerns, albeit potentially important ones,
such as ensuring proper citation of data and software or depositing
data and software in suitable repositories. Nevertheless,
we expect codecheckers to often recognise these issues and highlight them.

\textbf{How are failures during checks handled?}
We have no established process for the case
that a reproduction fails. This may be the sign of a pre-selection bias
of the examples, as so far all our CODECHECKs succeeded.
In case of a journal adopting CODECHECK for all submissions, the open 
question remains as what to do in this case will have to be answered.
At the start, we doubt publicly reporting failures (i.e.,~the code 
would not run) will increase overall reproducibility and recommend not
to share the outcome publicly but to share the latest draft of the 
certificate with the authors.
A negative comment in a CODECHECK certificate or a failed check does
not necessarily mean the paper or the science is bad
(cf. discussion on negative comments in \cite{everythinghertz123}).
Therefore, one must be careful as to the impact of acceptance by authors 
and the concerns of volunteers this approach, or a contrary one, has.
Private feedback for failures and public acknowledgement/encouragement
for successes seems to be the way to go.
% eLife's new model (see https://osf.io/9cftx/) will publish reports evantually, e.g.,~after preprint has been accepted elsewhere
\cite{Rosenthal2016b} discuss incentives for different actors around the
implementation of reproducibility. We see CODECHECK as one mean for
organisations to invest in reproducibility by creating incentives
until reproducible computations become the norm.

\textbf{Who will pay for the compute time?}
For those papers that take significant compute
time (days, not minutes), it is unclear who will pay for it.
The openness of CODECHECK certificates could help to establish community
peer pressure, or recognition, as submitting authors can be evaluated on their engagement
as codecheckers. This may lead to a balance of the shared load to reproduce
computations.
However, one must carfully consider the sustainability of rerunning computations and the
environmental impact large calculations, such as training machine learning models, have.
The challenges of computational resources can also quickly be reduced when
authors provide a small sample or a synthetic dataset together with the
expected outputs for this small example, which can equally well demonstrate the 
usefulness and validity of their workflow.

\textbf{What about my proprietary software and sensitive data?}
Authors can currently provide code that requires
proprietary software. Given the prevalence of proprietary software in
some disciplines, e.g MATLAB, this seems like a pragmatic choice.
However, it prohibits to benefit from open infrastructure for reproducibility
(cf.~\cite{konkol_publishing_2020,perkel_make_2019})
and requires the codechecker to have access to that particular software.
Non-open software also considerably hampers reuse, especially by researchers
from the global south. Therefore, allowing proprietary software is a compromise
that each implementation of CODECHECK should critically reconsider.

Solutions for proprietary and sensitive data exists.
Authors can provide synthetic data (cf. \cite{shannon_opening_2018}), some
data can effectively be redacted \cite{oloughlin_data_2015}, and publishers or
independent entities can provide infrastructure for sharing data and workflows
confidentially \cite{perignon_certify_2019} or with access to derived results
but not raw data \cite{shannon_opening_2018},
i.e.,~data enclaves \cite{foster_research_2018},
or domains of reproducibility \cite{harris_more_2017})

@@

\textbf{Who's got time for more peer review?} Agree.
Ignoring the complex community contract behind volunteered peer review,
the time spent with CODECHECKs likely reduces the individual's time to work
First, the \textbf{technical nature} of a CODECHECK can provide a more 
However, there are differences to peer review.
First, the technical nature of a CODECHECK can provide a more 
clearly defined scope and expectations and thereby time budget compared to conventional 
peer review. The authors are told what to provide and the codechecker
can be told when to stop, for example.
Second, the specific skill set allows to introduce different groups
into the peer review process. ECRs might be attracted to working with
groups to learn more about recent methods and reproducibility practices.
Research software experts (Research Software Engineers, RSE) who might
not regularly be involved in writing or reviewing papers could be 
interested to increase their connection with scholarly practices.
An extra codechecker may simplify the matchmaking an editor has to do
when identifying suitable reviewers for a submission, as technical and
topical expertise can be provided by different people, who must not even
be traditional peer reviewers. In fact, the codechecker could even be
a software professional employed by a publisher.
Third, recall that we expect CODECHECKs to be not anonymous and always
publicly deposited, which is common practice for some journals, but not
broadly practiced. The possibility to ask for clarification or more 
material/information can greatly reduce the digging required and thereby
increase \textbf{efficiency}.
Fourth, we also found that the focus on the mechanics of a
workflow and the ability to interact with the authors can make
reproductions very \textbf{educational}. It's a different role and as such could
be a welcome alternation/diversion for researchers to give back their 
time to the community.
With code and workflows, there also is a good chance that a codechecker's
feedback can have direct impact and lead to improvements of the author's 
work that comes with a rewarding experience or even tangible contributions.
While such impact is also part of idealistic peer review, it is
rather rare beyond paraphrased anonymous acknowledgement.

@@

As a principle, we do not require results to be exactly the same for a
CODECHECK to pass, simply that the code runs and generates the output
files that the author claims. Stochastic simulations mean that often
we cannot get exactly the same, or even the same versions of libraries
can generate outputs that differ by operating system
\cite{Gronenschild2012-pp}.
While reproducibility practices can mitigate some of these problems, e.g., by using a seed, the flexibility of the human judgement is still needed.

\textbf{Not revolutionary enough?}
CODECHECK's approach is to acknowledge the shortcomings around 
computational reproducibility and iteratively improve within the current
system of scholarly communication through journals.
It remains to be proved if this approach is welcomed broadly and if 
involving the traditional system of publishers helps to push the 
conversation along or not.
We have discussed stronger more drastic rules at length, such as only 
considering fully free and open source software, diamond Open Access journals or community-led no-fee journals,
and journals with collaborative/open peer review, but we eventually decided
against them on the level of the principles.
For the CODECHECK community process\footnote{\url{https://codecheck.org.uk/guide/community-process}}
and the volunteer codechecker community, these requirements are reconsidered.

We also have considered to require latest technologial solutions which can
support reproducibility (cf.~\cite{konkol_publishing_2020}), but have 
decided to focus on the human interface and the judgement of an experienced
researchers/developer as a more sustainable approach.
The structurelessness of the required data and software gives the needed
CODECHECK could be complemented by automated scoring (e.g.,~\cite{menke_rigor_2020}), yet automation and metrics bear new risks.
Note that the focus of the CODECHECK principles on code execution allows journals
and publishers to innovate on the financial model and peer review practices
at their own pace.
